---
title: 'Task 4: Naples Reef purple urchins'
author: "Larissa Neilson"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)
```

## Read in urchins dataset and convert column names to tidy format

```{r, echo = TRUE, message = FALSE, warning = FALSE}
urchins <- read.csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Create subset including only Naples Reef observations for purple urchins observed in January 2010

Begin by creating a subset with new month and year columns, and a new site column with the full name of the site, then make the desired subset

```{r, echo = TRUE, message = FALSE, warning = FALSE}
purple_urchins_dates <- urchins %>%
  mutate(date_new = mdy(date)) %>%
  mutate(year_new = year(date_new)) %>% 
  mutate(month_new =  month(date_new)) %>% 
  mutate(site_new = case_when(
    site == "NAPL" ~ "Naples Reef"))

purple_urchins_napl_2010 <- purple_urchins_dates %>%
  filter(common_name == "Purple Urchin", site_new == "Naples Reef",
         month_new == 1, year_new == 2010)
```

## Exploratory graphs of Naple Reef observations for purple urchins in January 2010

### Histogram of purple urchin size distribution by treatment

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(data = purple_urchins_napl_2010, aes(x = size)) +
  geom_histogram(bins = 8) +
  labs(x = "Size (cm)",
       y = "Count") +
  facet_wrap(~ treatment)
```

The size distribution for urchins under the annual treatment seems to be positively skewed while the control treatment distribution is slightly more normalized. I would not consider these samples to be normally distributed.

### QQ-plot of purple urchin size distribution by treatment

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(data = purple_urchins_napl_2010, aes(sample = size)) +
  geom_qq() +
  facet_wrap(~ treatment)
```

Clear positive skew in both treatments, but more so for the annual sample. I would not consider this data to be normally distributed.

## Create descriptive statistics table for both treatment groups

Then quickly calculate the actual difference in mean sizes vs. the percent difference of mean sizes between the two treatment groups.

```{r, echo = TRUE, message = FALSE, warning = FALSE}
purple_urchin_control_stats <- purple_urchins_napl_2010 %>%
  group_by(treatment) %>% 
  summarize(mean_size = mean(size),
            sd_size = sd(size),
            n = n(), .groups = 'drop')
kbl(purple_urchin_control_stats) %>% 
  kable_styling()

3.916667 - 3.725807
(3.916667 - 3.725807) / 3.725807
```

## Split treatments into different sample groups, and pull sample vectors

```{r, echo = TRUE, message = FALSE, warning = FALSE}
annual_sample <- purple_urchins_napl_2010 %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size)

control_sample <- purple_urchins_napl_2010 %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size)
```

## Run two-sample t-test with the two treatment vectors as inputs

```{r, echo = TRUE, message = FALSE, warning = FALSE}
t.test(annual_sample, control_sample)
```
The means of these two samples (annual and control treatments) are quite similar, with only a five percent difference between them. This  tells me that it is somewhat likely that they could have been drawn from populations with the same mean.

The p-value is a probability that tells me that if these samples are drawn from populations with the same mean size, then there is approximately a 25% chance that I could have taken two random samples from those populations and found samples at least as different as the samples that I have here.

This tells me that it is possible that I could have retrieved my samples from populations with the same mean, but it is uncertain.

Further observations are necessary with sample sizes larger than n = 60, 62 to make a more concrete conclusion of the data.
