---
title: "Task 3 & 4"
author: "Larissa Neilson"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Read in urchins dataset and make column names coding friendly

```{r, echo = TRUE, message = FALSE, warning = FALSE}
urchins <- read.csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Exploratory data visualization

Histogram of urchin size distribution

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(data = urchins, aes(x = size)) +
  geom_histogram(bins = 15) +
  labs(x = "Size (cm)",
       y = "Count") +
  facet_wrap(~ common_name)
```

The purple urchin size distribution is positively skewed and narrow, while the red urchin size distribution is slightly positively skewed with a broader distribution. I don't think I would consider this data to be normally distributed.

Boxplot of urchin size distribution 

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(data = urchins, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = NULL,
       y = "Size (cm)")
```

The average size of purple urchins is a little above 4 cm, while the red urchins are generally much bigger, at ~7 cm.

# Create a descriptive statistics table of urchin data

```{r, echo = TRUE, message = FALSE, warning = FALSE}
urchin_stats <- urchins %>%
  group_by(common_name) %>% 
  summarize(mean_size = mean(size),
            sd_size = sd(size),
            n = n())
  kbl(urchin_stats) %>% 
  kable_styling()
```
The first thing I notice about these statistics is that the sample size for purple urchins is much larger than for red urchins, which is probably why we see a smaller standard deviation for purple urchins (1.18) versus red urchins (1.89).

## Exploring purple urchin counts by site and year

```{r, echo = TRUE, message = FALSE, warning = FALSE}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date_new = ymd(date)) %>%
  mutate(year_new = year(date_new)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))

```

Purple urchins summary table

```{r, echo = TRUE, message = FALSE, warning = FALSE}
purple_urchin_stats <- purple_urchins %>% 
  group_by(site_full, year_new) %>% 
  count(common_name)
kbl(purple_urchin_stats) %>% 
  kable_styling()
```

## Exploratory column graph of purple urchins

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(data = purple_urchin_stats, aes(x = year_new, y = n)) +
         geom_col(aes(fill = site_full,
                      show.legend = FALSE)) +
  labs(x = "Year", y = " Urchin Count") +
  scale_x_continuous(breaks = c(2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012)) +
  coord_flip() +
  facet_wrap(~site_full) +
  theme(legend.position = "none")
```

